---
title: "Spatial Regression"

date: <h4 style="font-style:normal">Aug 31, 2021</h4>
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
---


<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: normal;
}

h1.title {
  font-weight: bold;
  font-family: Arial;  
}

h2.title {
  font-family: Arial;  
}

</style>


<style type="text/css">
#TOC {
  font-size: 13px;
  font-family: Arial;
}
</style>


\

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

# Study set up

READ THIS WITH THE READINGS:
https://psu-spatial.github.io/Stat462-2021/index.html#S.Tutorial.1A
and for the spatial, that lab 7 from 364

## Load libraries

```{r, message=FALSE, warning=FALSE}
# if a library doesn't exist, then to download, go to the packages menu (next to help) and click install

# Upon downloading the .rmd file, I attempted to run this first code chunk and found that many of the libraries weren't installed.
#I needed to install each library before downloading it via this code

## pacman library is cool check it out


rm(list=ls())
library(sp)
library(sf)
library(raster)
library(tidycensus)
library(tidyverse)
library(plotly)
library(olsrr)
library(tmap)
library(RColorBrewer)
library("lubridate")
library(dplyr)
library(tmap)
library(olsrr)
library(skimr)
```

## Read in data

This data was accessed from.....

```{r}
# dot . means current folder - and you don't need a file extension

# this code chunk uses "st_read" to read in the features from the initial tornado point shapefile
# the table/spreadsheet of data is saved to a variable "T_data"
T_data <- st_read(dsn="./1950-2019-torn-initpoint",
                  layer="1950-2019-torn-initpoint")

```


## Initial set up

```{r}
# make the date
# set the values in the date column to the variable "T_data$date"
# as.Date command converts the character data in the column to represent calendar dates
# calendar date is in the format year, month, date

T_data$date <- as.Date(T_data$date, format="%Y-%m-%d")
```

```{r}

# strsplit splits each character time string into a "list" (advent calendar) 
# of the hour, minutes and seconds for each row 
Fulltimes <- strsplit(T_data$time,":")


# Now we use the lapply function to make a new list with just the first one
# and we turn it back into just a normal row of numbers
T_data$hours   <-as.numeric(unlist(lapply(Fulltimes, "[",1)))
T_data$minutes <-as.numeric(unlist(lapply(Fulltimes, "[",2)))
T_data$seconds <-as.numeric(unlist(lapply(Fulltimes, "[",3)))
T_data$DecimalHour <- T_data$hours + (T_data$minutes/60)
```

```{r}
# mag < 0 has been set to missing
T_data$mag[T_data$mag < 0 ] <- NA
```

## Time of day

```{r}
# Add in time of day
T_data$TimeOfDay <- NA
T_data$TimeOfDay[T_data$hours %in% c(0:5)] <- "Sleep_0_5"
T_data$TimeOfDay[T_data$hours %in% c(6:11)]    <- "Morning_6_11"
T_data$TimeOfDay[T_data$hours %in% c(12:17)]    <- "Afternoon_12_17"
T_data$TimeOfDay[T_data$hours %in% c(18:23)] <- "Evening_18_23"

tmap_mode("plot") # view
#qtm(T_data,dots.col="TimeOfDay")

```

```{r}
# The ~ means "is a function of" e.g. y variable goes first

# wrong plot command for categorical data
#plot(T_data$fat, T_data$TimeOfDay,
#     ylab="Average number of fatalities in that hour slot",
#     xlab="Time of day",
#     type="p",pch=16,cex=.1,col=rgb(0,0,0,.5))
```


```{r}
#skim(T_data)
```

```{r}
###  two equals means "is equal to??"
## != not equal
## < > MAKE TEST DATA THAT IS SPATIAL
 Catdata <- T_data[T_data$mag == 4,]
Catdata <- Catdata[Catdata$loss < 1000000,]
Catdata.sf <- st_as_sf(Catdata,coords=c("elon","elat"),crs=4236)
```


```{r}
model1 <- lm(loss ~ slat, data=Catdata.sf)
ols_regress(model1)
plot(loss ~ slat, data=Catdata.sf)
```



```{r}
model2 <- lm(loss ~ slat+TimeOfDay+fat, data=Catdata.sf)
ols_regress(model2)
```

# comparing two models

```{r}
# to compare
anova(model1,model2)
```


```{r}
# lower is better
AIC(model1,model2)
```
# seeing if our favourite model meets 

THE 4 ASSUMPTIONS - GO FIND THESE AND WRITE THEM DOWN
(LINE)
https://psu-spatial.github.io/Stat462-2021/index.html#S.Tutorial.7C


```{r}
ols_plot_resid_fit(model1)

```


```{r}
ols_plot_resid_stud(model1)

```

```{r}
ols_plot_resid_qq(model1)

```

```{r}
ols_plot_resid_hist(model1)

```

```{r}
ols_test_normality(model1)

```

```{r}
ols_plot_cooksd_bar(model1)

```

```{r}
ols_plot_resid_lev(model1)

```

# Spatial residuals

```{r}
Catdata.sf$model1.Residuals <- residuals(model1)

```

